---
title: "DNA boxplot and depth plots, both metagenome and amplicon data"
author: "Nathan R. Geraldi"
date: "May 21, 2019"
output: github_document
---

set table options
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## libraries
```{r libraries, message=FALSE, warning=FALSE}
library(RColorBrewer)
library(tidyverse) 
library(broom)
#library(edgeR)
library(corrplot)

#  install.packages("devtools")
#  devtools::install_github("grunwaldlab/metacoder")

```

## functions
```{r functions, message=FALSE, warning=FALSE}
error.bar <- function(x, y, upper, lower=upper, length=0.04, ...){
  if(length(x) != length(y) | length(y) !=length(lower) | length(lower) != length(upper))
    stop("vectors must be same length")
  arrows(x+upper,y, x-lower, y, angle=90, code=3, length=length, lwd= 2, ...)
}

```


## define universal variables
```{r define_universal}
# stud_pat<-"NA"  # matches study specific title from pipe (begining of files).
dir<-"/Users/nathangeraldi/Dropbox/"
out_file<-"Documents/KAUST/eDNA/R/pipe_summary"
# export  to project folder
export_file<-"Documents/KAUST/eDNA/R/csv/"
# for plots
export_cor_plot<-"Documents/KAUST/eDNA/DMAP/R/plots/all_amplicon/"

# folder where scripts are
wd_now<-paste(dir,"/Documents/KAUST/eDNA/R/projects/Global_ocean_genome_analysis/",sep="")

```

## import misc
```{r import_misc}
## taxa conversions to compare net data
taxa_DNA_net<-openxlsx::read.xlsx(xlsxFile=paste(dir,"Global_databases/DMAP/tow_biomass/taxa_table.xlsx", sep=""), sheet=1)
taxa_net_DNA<-openxlsx::read.xlsx(xlsxFile=paste(dir,"Global_databases/DMAP/tow_biomass/taxa_table.xlsx", sep=""), sheet=2)

# get ocean data  - used in "tidy 18s"
geo<-read.table(paste(dir,"Documents/KAUST/eDNA/DMAP/R/CSV/Global_layers_oct18.csv",sep=""), sep=",",header=T)
geo1<-geo[,c(1,22)]# get lohg and ocean   names(geo)
geo_ll<-geo[,c(2,3,6,7,22,13)]#   column 3 is station
geo_ll<-geo_ll[!duplicated(geo_ll$Station),]
geo_ll<-geo_ll[,-2]
#
```



## import amplicon data
this was from amplicon_phylums4.R 
```{r amplicon_import}

##     get TARA data
# use this script to import and clean data
source(paste(wd_now,"amplicon_stats_tara_source.R", sep="")  )
pat<-c("Annelida","Arthropoda","Chordata","Cnidaria","Ctenophora","Echinodermata","Gastrotricha","Mollusca",
       "Nematoda","Orthonectida","Platyhelminthes","Porifera","Rotifera","Sipuncula","Tardigrada","Urochordata","Chaetognatha",
       "Craniata","Nemertea","Bryozoa","Hemichordata","Brachiopoda","Entoprocta","Cephalochordata","Entoprocta") 
pat<-paste(pat, collapse = "|")
### make ocean basins
sdat<- dat_otu_rare$read_low_sp  
ob_cat<-data.frame(unique(sdat$Locality.2))
ob_cat$ocean_basin<-"Mediterranean Sea"
ob_cat$ocean_basin[2]<-"Red Sea"
ob_cat$ocean_basin[c(3,4,5)]<-"Indian Ocean"
ob_cat$ocean_basin[c(7,8,10)]<-"Atlantic"
ob_cat$ocean_basin[9]<-"Southern Ocean"
ob_cat$ocean_basin[c(11:13)]<-"Pacific"
ob_cat$ocean_basin[14]<-"Mediterranean Sea"
names(ob_cat)[1]<-"Locality.2"


###### TARA -  tdat   sum to phylum 
#  names(read_low)
tdat1<- dat2 %>%   # 247160    names(dat2)
  filter(pid>90) %>%  # only greater than 90 PID
  group_by_at(vars(Sample.ID,Station:filter_size,OTU.LEVEL.SHANNON.DIVERSITY_from.Vargas:Locality.4,filter_size_cat,Phylum)) %>% # was sp_num in stats_source
  summarize(reads=sum(reads),reads_rarified=sum(reads_rarified), reads_percent=sum(reads_percent), rich=length(which(percent.per.metazoan.reads>0))) %>% 
  filter(complete.cases(Phylum)) %>% 
  filter(Phylum != "")  %>% 
  left_join(ob_cat) %>% 
  group_by(Sample.ID) %>% 
  mutate(new_tot_reads=sum(reads)) %>% 
  ungroup() %>% 
  mutate(new_prop_reads=reads/new_tot_reads)
#  length(unique(tdat1$Station))
tdat2<-data.frame(tdat1)

## tara sum to order
tdat_order<- dat2 %>%   # 247160    names(dat2)
  filter(pid>90) %>%  # only greater than 90 PID
  group_by_at(vars(Sample.ID,Station:filter_size,OTU.LEVEL.SHANNON.DIVERSITY_from.Vargas:Locality.4,filter_size_cat,Phylum:Order)) %>% # was sp_num in stats_source
  summarize(reads=sum(reads),reads_rarified=sum(reads_rarified), reads_percent=sum(reads_percent), rich=length(which(percent.per.metazoan.reads>0))) %>% 
  filter(complete.cases(Order)) %>% 
  filter(Order != "")  %>% 
  left_join(ob_cat) %>% 
  group_by(Sample.ID) %>% 
  mutate(new_tot_reads=sum(reads)) %>% 
  ungroup() %>% 
  mutate(new_prop_reads=reads/new_tot_reads)
##############################################################################################
######   Mal data          ######################################################################################

source(paste(wd_now,"amplicon_stats_mal_source.R", sep="") )
names(dat)

# summarize by phylum
mdat1<- dat %>%   # 247160    names(mdat1)
  #filter(DNAID=="DNA") %>% 
  filter(pid_col>90) %>%  # only greater than 90 PID
  group_by_at(vars(Sample.ID,DNAID,Station:filter_size,Phylum)) %>% # was sp_num in stats_source
  summarize(rich=length(which(reads_percent>0)),reads=sum(reads),reads_rarified=sum(reads_rarified), reads_percent=sum(reads_percent) ) %>% 
  filter(complete.cases(Phylum)) %>% 
  filter(Phylum != "") %>% 
  group_by(Sample.ID) %>% 
  mutate(new_tot_reads=sum(reads)) %>% 
  ungroup() %>% 
  mutate(new_prop_reads=reads/new_tot_reads) %>% 
  mutate(Depth_cat=cut(Depth, breaks=c(-Inf,11, 200, 1000, 2000, 3910, Inf), labels=c("<10m","10-200m","200-1000m","1000-2000m","2000-3910m","3995-4020m")))
#
mdat2<-data.frame(mdat1)  # class(mdat1)

mdat_order<- dat %>%   # 247160    names(mdat1)
  #filter(DNAID=="DNA") %>% 
  filter(pid_col>90) %>%  # only greater than 90 PID
  group_by_at(vars(Sample.ID,DNAID,Station:filter_size,Phylum:Order)) %>% # was sp_num in stats_source
  summarize(rich=length(which(reads_percent>0)),reads=sum(reads),reads_rarified=sum(reads_rarified), reads_percent=sum(reads_percent) ) %>% 
  filter(complete.cases(Order)) %>% 
  filter(Order != "") %>% 
  group_by(Sample.ID) %>% 
  mutate(new_tot_reads=sum(reads)) %>% 
  ungroup() %>% 
  mutate(new_prop_reads=reads/new_tot_reads) %>% 
  mutate(Depth_cat=cut(Depth, breaks=c(-Inf,11, 200, 1000, 2000, 3910, Inf), labels=c("<10m","10-200m","200-1000m","1000-2000m","2000-3910m","3995-4020m")))



```

## tidy 18s phyl
```{r tidy_18S}
## set for both 18s and genome !!!!!!
top10all<-c("Arthropoda","Cnidaria","Nematoda","Chordata","Ctenophora","Annelida","Placozoa","Mollusca", "Chaetognatha", "Tardigrada","Platyhelminthes")


### for phyla
tara<-tdat2 %>%    #  names(tdat2)   unique(tdat10$Phylum)
  select(Sample.ID,Cruise.x,Depth_region,filter_size_cat,Latitude,ocean_basin,Phylum,new_prop_reads, rich) %>% 
  mutate(DNAID="DNA") 

dat11<-mdat2 %>%    #  names(mdat2) names(geo1)    !!!!!!!!!
  mutate(Cruise.x="Malaspina") %>% 
  select(Sample.ID,Cruise.x,DNAID,Depth_cat,Latitude,Phylum,new_prop_reads,rich) %>% 
  bind_rows(tara) %>% #  add TARA
  left_join(geo_ll) %>% 
  mutate(ocean=as.character(ocean)) %>%  
  mutate(ocean2=ifelse(is.na(ocean), ocean_basin, ocean)) %>% 
  mutate(ocean2=replace(ocean2, ocean2=="Antarctic","Southern Ocean")) %>% 
  mutate(Latitude_abs=abs(Latitude)) %>% 
  mutate(Latitude_cat=cut(Latitude_abs, breaks=c(-Inf,23.5,35, Inf), 
                        labels=c("Tropical","Subtropical","Temperate")))  #   catagory
####    names(dat11)
mess<- dat11 %>% # names(dat11)  unique(dat11$Phylum)
  filter(DNAID=="DNA") %>% 
  group_by(Phylum) %>% 
  summarize(reads=mean(new_prop_reads,na.rm = T), med=median(new_prop_reads,na.rm = T), mean_rich=mean(rich,na.rm = T) , med_rich=median(rich,na.rm = T)) %>% 
  arrange(desc(reads)) %>% # old but left in
  arrange(desc(mean_rich))
  
##  quick look at reads
tss<-sum(mess$reads)
tp10<-sum(mess$reads[1:10])/tss # first 10 have 0.999 % or reads
tpat<-mess$Phylum[1:15] ## top 10 phylum
## at richness
tsr<-sum(mess$mean_rich) #

# top ten of both-12 phylua  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
top10pat_rich<-mess$Phylum[1:10]
  ###  limit datat to top 10 overall
Amp_10 <- dat11 %>% 
  filter(grepl(paste(top10pat_rich, collapse="|"), Phylum)) %>% 
  mutate(Phylum=factor(Phylum, levels=top10pat_rich)) 

top10pat18s<-top10pat_rich


```


## tidy 18s order
```{r tidy_18S}

### for phyla
tara_order<-tdat_order %>%    #  names(tdat2)   unique(tdat10$Phylum)
  select(Sample.ID,Cruise.x,Depth_region,filter_size_cat,Latitude,ocean_basin,Phylum:Order,new_prop_reads, rich) %>% 
  mutate(DNAID="DNA") 

dat1_order<-mdat_order %>%    #  names(mdat2) names(geo1)    !!!!!!!!!
  mutate(Cruise.x="Malaspina") %>% 
  select(Sample.ID,Cruise.x,DNAID,Depth_cat,Latitude,Phylum:Order,new_prop_reads,rich) %>% 
  bind_rows(tara_order) %>% #  add TARA
  left_join(geo_ll) %>% 
  mutate(ocean=as.character(ocean)) %>%  
  mutate(ocean2=ifelse(is.na(ocean), ocean_basin, ocean)) %>% 
  mutate(ocean2=replace(ocean2, ocean2=="Antarctic","Southern Ocean")) %>% 
  mutate(Latitude_abs=abs(Latitude)) %>% 
  mutate(Latitude_cat=cut(Latitude_abs, breaks=c(-Inf,23.5,35, Inf), 
                        labels=c("Tropical","Subtropical","Temperate")))  #   catagory
####    names(dat11)
mess<- dat1_order %>% # names(dat11)  unique(dat11$Phylum)
  filter(DNAID=="DNA") %>% 
  group_by(Phylum,Class,Order) %>% 
  summarize(reads=mean(new_prop_reads,na.rm = T), med=median(new_prop_reads,na.rm = T), mean_rich=mean(rich,na.rm = T) , med_rich=median(rich,na.rm = T)) %>% 
  arrange(desc(reads)) %>% # old but left in
  arrange(desc(mean_rich))
  
##  quick look at reads
tss<-sum(mess$reads)
tp10<-sum(mess$reads[1:10])/tss # first 10 have 0.999 % or reads
tpat<-mess$Phylum[1:15] ## top 10 phylum
## at richness
tsr<-sum(mess$mean_rich) #

# top nine of both-12 phylua
top20pat<-mess$Order[1:18]  # 18 has => 1 mean rich per sample
  ###  limit datat to top 10 overall
Amp_Order_20 <- dat1_order %>% 
  filter(grepl(paste(top20pat, collapse="|"), Order)) %>% 
  mutate(Order=factor(Order, levels=top20pat)) 
Amp_Order_20 <-data.frame(Amp_Order_20)

top20pat18s_order<-top20pat

### get colors so order/class is same as phylum

Amp_Order_20_colors <- Amp_Order_20 %>% 
  select(Phylum:Order) %>% 
  distinct_all() %>% 
  mutate(Order=factor(Order, levels=top20pat18s_order)) %>% 
  arrange(Order)

```


## metagenome
```{r metagenome}

# genome data from DMAP        
dat_genome<-data.table::fread(file = paste(dir, 'Documents/KAUST/eDNA/DMAP/R/CSV/DMAP_biomass_apr19.csv',sep=""), sep = ',', header = TRUE)
##   add column differentiat malaspina profile and deep
dat_genome1<- dat_genome %>%   # names(dat_genome)      
  mutate(pid=forcats::fct_recode(pid, "PID_50" ="pid50","PID_70" ="pid70","PID_90" ="pid90" )) %>% 
  mutate(Depth = ifelse(Cruise == "TARA", Depth.env, Depth)) %>% 
  group_by_at(vars(unique_ID:pid, Latitude,Longitude,Depth,phylum)) %>% # was sp_num in stats_source
  summarize(reads=sum(reads)) %>% 
  ungroup() %>% 
  group_by_at(vars(unique_ID:pid)) %>% 
  mutate(reads_sample=sum(reads)) %>% 
  ungroup() %>% 
  mutate(reads_percent=reads/reads_sample) %>% 
  dplyr::rename(Phylum=phylum) 
  # unique(dat_genome1$pid)
  ## get oceans, but need to remove some samples without lat/lon
 dat_genome_na<- dat_genome1 %>% # remove then bind back in after join-erro with na's
    filter(is.na(Latitude))

 dat_genome2<- dat_genome1 %>% # names(dat_genome2)
    filter(complete.cases(Latitude)) %>% 
    fuzzyjoin::geo_left_join(geo_ll, max_dist = 30)  %>% #  to get oceans, dist in miles
    select(-Latitude.y,  -Longitude.y) %>% 
    dplyr::rename(Latitude=Latitude.x , Longitude=Longitude.x)  %>% 
    mutate(ocean=as.character(ocean)) %>%  #  unique(dat_genome2$ocean)
    mutate(ocean=replace(ocean, ocean=="Antarctic","Southern Ocean")) %>% 
    mutate(Latitude_abs=abs(Latitude)) %>% 
    mutate(Latitude_cat=cut(Latitude_abs, breaks=c(-Inf,23.5,35, Inf), 
                        labels=c("Tropical","Subtropical","Temperate"))) %>% 
   bind_rows(dat_genome_na) %>% 
   # remove NaN- only for samples with no hits
   filter(!reads_percent=="NaN") 

mes<- dat_genome %>% 
  filter(Cruise=="TARA", phylum=="Mollusca")
 #########
  #       get filter catagories
filcat<-data.frame(unique(dat_genome2$filter_size))
colnames(filcat)[1] <- "filter_size"
filcat$filter_size_cat<-c("0.2-3","1-20","1-20","0.2-3","0.2-3","1-20","0.1-0.22","0.2-3","0.2-3","0.2-3","")
dat_genome2<-left_join(dat_genome2, filcat, by="filter_size")   #  unique(dat_genome2$filter_size_cat)
##  get depth catagories
dat_genome2$Depth_cat <- cut(dat_genome2$Depth, breaks=c(-Inf,11, 200, 1000, 2000, 3910, Inf), 
                      labels=c("<10m","10-200m","200-1000m","1000-2000m","2000-3910m","3995-4020m"))
##
# get order of phyla most to least
top10patMG <-dat_genome2 %>% 
  group_by(Phylum) %>% 
  summarize(m=mean(reads_percent),sum=sum(reads_percent)) %>% 
  arrange(desc(m)) %>% 
  top_n(10) %>% 
  pull(Phylum) 
####     
# limit data to top 10 and search
dat_genome3<-dat_genome2 %>%   # unique(sum1$fac)  names(dat_phyl)  names(dat_genome2)
  filter(Phylum %in% top10patMG) %>%   # limit to top ten phyla    top10patMG  or top10all
  mutate(Phylum=factor(Phylum, levels=top10patMG))  %>% 
  # limit gene domain    !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(gene_domain=="ec4")  %>%   # unique(dat_genome2$gene_domain)
   # limit to target DMAP filter
  filter(pid=="PID_70")  #  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

dat_genome3<-data.frame(dat_genome3)

mes<-dat_genome3 
  
```


## net data
Remove ? -- Yes
```{r net_data}
# overall biomass - don't need
#mal_net_bio<-read.csv("/Users/geraldn/Dropbox/Global_databases/DMAP/tow_biomassZooplankton #Biomass Malaspina.csv")

# abundance from Mariluz
mal_net_abun<-openxlsx::read.xlsx(xlsxFile="/Users/geraldn/Dropbox/Global_databases/DMAP/tow_biomass/Zooplankton_Malaspina_abund.xlsx", sheet=1)
mal_net_abun_depth_convert<-openxlsx::read.xlsx(xlsxFile="/Users/geraldn/Dropbox/Global_databases/DMAP/tow_biomass/Zooplankton_Malaspina_abund.xlsx", sheet=2)

## biomass from Santiago, this sheet also has abundance
mal_net_biomass<-openxlsx::read.xlsx(xlsxFile="/Users/geraldn/Dropbox/Global_databases/DMAP/tow_biomass/tow_biomass_ABDCIA_all_MALASPINA.xlsx", sheet=1)

# biomass      names(mal_net_biomass)
mal_net_biomass2 <- mal_net_biomass %>% 
  dplyr::rename( "Station"="Estación","strata"  ="Estrato","Depth" ="Profundidad.(m)", "Chaetognatha"="Bio.Chaetognatha", "Copepoda"="Bio.Copepoda", "Euphausiids.Like"="Bio.Euphausiids.Like", "Gelatinous"="Bio.Gelatinous", "Other.Mesoz."="Bio.Other.Mesoz.") %>% 
  separate(Depth, into=c("min_depth","max_depth"), sep="-", remove=TRUE, convert=TRUE) %>% 
  gather(key=taxa, value=biomass, Chaetognatha:Other.Mesoz.) %>% 
  mutate(biomass_percent=biomass/Bio.total) %>% 
  mutate(taxa_j=taxa) %>% 
  mutate(taxa_j=replace(taxa_j, taxa_j=="Euphausiids.Like","Arthropoda")) %>% 
  mutate(taxa_j=replace(taxa_j, taxa_j=="Copepoda","Arthropoda"))

# abudnance   #  names(mal_net_abun2)  mal_net_abun_depth_convert
mal_net_abun2 <- mal_net_abun %>% 
  gather(key=taxa, value=abundance_percent, Jellyfish:Meroplankton.Larvae)  %>% 
  left_join(mal_net_abun_depth_convert) %>% 
  left_join(taxa_net_DNA, by=c("taxa" = "abundance_taxa")) #  %>% 
# ???

```




## box 18 richness
plot richenss of phyla for 18s

```{r barplots_per}

#####
#  lfil<-c(".5-5","5-20","20-200","180-2000")
n<-16
cc1<-rainbow(n)
cc1<-c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1','#35978f','#01665e','#003c30', "black")
space<-c(1:10) #
cc1<-cc1[1:length(space)]
yylim<-c(0,50)
ddd2<-3 # set divisor of yylim to adjust ylims
ddd<-6 # divisor for yylim3

tttlx<-10  ## location of plot label
tttly<-45
tttletx<-0.5    # location of letter
tttlety<-48
ypp<--2 # y of x axis labels
bxlwd<-1.5  # bos line width
mean_pch<-19 # pch of mean 
mean_cex<-1.3 #size o mean circle 


############################################################
#### plot    names(dat1)
pdf(paste(dir, export_cor_plot,'boxplot_18s_richness_phyl_order.pdf',sep = ""), width=10, height=10 )

layout.matrix <- matrix( c(1,2,3,4,5,c(6:10) ), nrow = 5, ncol = 2, byrow = FALSE)

layout(mat = layout.matrix,
       heights = c(1,1,1,1,1), # Heights of the two rows
       widths = c(1, 1.7)) # Widths of the two columns
#  layout.show(10)



par(mar=c(.2,1,1,2), oma=c(8,3,0,0), xpd = NA) 
########

############################################################
############################################################
# start column 1  Amplicon data
############################################################
####    TARA
rr<-c(8) # response variable in this col  !!!!
r<-1  ### number of levels
# tara ocean surface - comparable to mala
pdat<-Amp_10 %>%   ## unique(dat11$Cruise.x)  names(pdat) unique(pdat$filter_size_cat) 
  filter(Cruise.x=="TARA") %>% 
  filter(Depth_region=="SRF") %>% 
  filter(filter_size_cat==".5-5" )
j<-rr
fac1<-pdat$Phylum
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, ann=FALSE)   # , log="y"
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Tara surface", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="A", cex=1)    
#mtext(text="b",side=1, line=-13.0,cex=cx, adj=.08, font=2)
############################################################
# malaspina surface  2nd plot
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("M",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
 j<-rr
fac1<-pdat$Phylum
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim/ddd, boxlwd = bxlwd, ann=FALSE)   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly/ddd, label="Malaspina surface", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety/ddd, label="B", cex=1)  
###################################################################################
#########################################
### malaspina profile - DNA then RNA     names(dat11)
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(!(grepl("deep",Sample.ID) | grepl("M",Sample.ID)))  # names(dat1)
fac3<-factor(pdat$DNAID)
ll<-levels(fac3)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-pdat1[,j]
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim/ddd, boxlwd = bxlwd, ann=FALSE)   # , log="y"
    if (i==3) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=top10pat,tck=-0.01, cex=.9)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-c("Malaspina profile","Malaspina Profile  RNA")
    text(tttlx,tttly/ddd, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("C","D")
    text(tttletx,tttlety/ddd, label=ttlet[i], cex=1)  
  }
}

############################################################
# malaspina deep
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("deep",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
j<-rr
fac1<-pdat$Phylum
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim/ddd2, boxlwd = bxlwd, ann=FALSE)   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
##   !! only when need x axis labels
text(x=space, y=rep(ypp,10), labels=top10pat_rich, adj=1, srt=60, xpd = NA)  # !!!!!
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
#  med<-tapply(resp,list(fac1),median,na.rm=T)
#  er<-tapply(resp,list(fac1),plotrix::std.error,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly/ddd2, label="Malaspina deep", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety/ddd2, label="E", cex=1)  
############################################################




########################################################################################################################
########################################################################################################################
############################################################
# start column 2 by smaller lineage than phyla -- order?
############################################################
# names(Amp_Order_20)
space<-c(1:length(top20pat18s_order))
yylim<-c(0,20)
ddd<-4
dd5<-4 ## adjust first mala
tttlety<-19
tttletx<-0.2
ypp<--0.5 # y level of x axis labels
## quick match colors to phylums
## ## cc1      Amp_Order_20_colors
or_col<- top10pat18s %>% 
  bind_cols(cc1[1:10]) %>% 
  rename_all(funs(c("Phylum","color1")))
cc1 <- Amp_Order_20_colors %>% 
  left_join(or_col) %>% 
  pull(color1)

# remve\\!!!
#Amp_Order_20<-Amp_Order_20 %>% 
#  left_join(or_col) %>% 
#  replace_na(list(color = "black")) 
#Amp_Order_20<-data.frame(Amp_Order_20)
  

####    TARA
rr<-c(10) # response variable in this col  !!!!
r<-1  ### number of levels
# tara ocean surface - comparable to mala
pdat<-Amp_Order_20 %>%   ## unique(dat11$Cruise.x)  names(pdat) unique(pdat$filter_size_cat) 
  filter(Cruise.x=="TARA") %>% 
  filter(Depth_region=="SRF") %>% 
  filter(filter_size_cat==".5-5" )
j<-rr
fac1<-pdat$Order
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F, col="white",axes=F, at=space, border=cc1, ylim=yylim, boxlwd = bxlwd, ann=FALSE)   # , log="y"
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
#text(tttlx,tttly, label="Tara surface", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="F", cex=1)    
#mtext(text="b",side=1, line=-13.0,cex=cx, adj=.08, font=2)
############################################################
# malaspina surface  2nd plot
pdat<-Amp_Order_20 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("M",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
 j<-rr
fac1<-pdat$Order
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F, at=space, border=cc1, ylim=yylim/dd5, boxlwd = bxlwd, ann=FALSE)   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
#text(tttlx,tttly/ddd, label="Malaspina surface", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety/dd5, label="G", cex=1)  
###################################################################################
#########################################
### malaspina profile - DNA then RNA     names(dat11)
pdat<-Amp_Order_20 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(!(grepl("deep",Sample.ID) | grepl("M",Sample.ID)))  # names(dat1)
fac3<-factor(pdat$DNAID)
ll<-levels(fac3)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Order
    #fac2<-dat11$filter_crit_cat
    resp<-pdat1[,j]
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim/ddd, boxlwd = bxlwd, ann=FALSE)   # , log="y"
    if (i==3) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=top10pat,tck=-0.01, cex=.9)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-c("Malaspina profile","Malaspina Profile  RNA")
   # text(tttlx,tttly/ddd, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("H","I")
    text(tttletx,tttlety/ddd, label=ttlet[i], cex=1)  
  }
}

############################################################
# malaspina deep
pdat<-Amp_Order_20 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("deep",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
j<-rr
fac1<-pdat$Order
resp<-pdat[,j]
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim/dd5, boxlwd = bxlwd, ann=FALSE)   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
##   !! only when need x axis labels
text(x=space, y=rep(ypp,length(space)), labels=top20pat18s_order, adj=1, srt=60, xpd = NA)  # !!!!!
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
#  med<-tapply(resp,list(fac1),median,na.rm=T)
#  er<-tapply(resp,list(fac1),plotrix::std.error,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
#text(tttlx,tttly/ddd2, label="Malaspina deep", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety/dd5, label="J", cex=1)  
############################################################


#####################################################################################
# at end add titlas
cqq<-.8
l11<-.2
mtext("Richness (metabarcode)", 2,las=0,line=1.7, cex=cqq, at=.5 ,outer=TRUE)
mtext("Phylum", 1,las=0,line=6.5, cex=cqq, at=.15 ,outer=TRUE)
mtext("Order", 1,las=0,line=6.5, cex=cqq, at=.65 ,outer=TRUE)
#  !!!!!!!!!!!!!
dev.off()

```








## box 18 and gen reads
plot percents of phyla for both 18S and metagenomes
NOT USED IN MANUSCIPT !!!!!!
```{r barplots_per}
lfil<-c(".5-5","5-20","20-200","180-2000")
n<-16
cc1<-rainbow(n)
cc1<-c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1','#35978f','#01665e','#003c30', "black")
space<-c(1:11) #
cc1<-cc1[1:length(space)]
yylim<-c(1,100)
tttlx<-11  ## location of plot label
tttly<-45
tttletx<-11    # location iof letter
tttlety<-80
ypp<-0.65
bxlwd<-1.5  # bos line width
mean_pch<-19 # pch of mean 
mean_cex<-1.3 #size o mean circle 


############################################################
#### plot    names(dat1)
par(mfcol=c(6,3), mar=c(.2,1,1,2), oma=c(8,3,0,0), xpd = NA) 
########

############################################################
############################################################
# start column 1  Amplicon data
############################################################
####    TARA
rr<-c(7) # response in col  !!!!
r<-1  ### number of levels
# tara ocean surface - comparable to mala
pdat<-Amp_10 %>%   ## unique(dat11$Cruise.x)  names(pdat) unique(pdat$filter_size_cat) 
  filter(Cruise.x=="TARA") %>% 
  filter(Depth_region=="SRF") %>% 
  filter(filter_size_cat==".5-5" )
j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Tara surface 18S", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="A", cex=1)    
#mtext(text="b",side=1, line=-13.0,cex=cx, adj=.08, font=2)
############################################################
# malaspina surface  2nd plot
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("M",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
 j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina surface 18S", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="B", cex=1)  
###################################################################################
#########################################
### malaspina profile - DNA then RNA     names(dat11)
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(!(grepl("deep",Sample.ID) | grepl("M",Sample.ID)))  # names(dat1)
fac3<-factor(pdat$DNAID)
ll<-levels(fac3)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-((pdat1[,j])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   # , log="y"
    if (i==3) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=top10pat,tck=-0.01, cex=.9)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-c("Profile 18S","Profile 18S RNA")
    text(tttlx,tttly, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("C","D")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
  }
}

############################################################
# malaspina deep
pdat<-Amp_10 %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(grepl("deep",Sample.ID)) %>% 
  filter(DNAID=="DNA") 
j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
#  med<-tapply(resp,list(fac1),median,na.rm=T)
#  er<-tapply(resp,list(fac1),plotrix::std.error,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina deep 18S", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="E", cex=1)  
############################################################
# ALL  18S  #
j<-rr
pdat<-Amp_10 %>% 
  filter(DNAID=="DNA") 
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd =bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=2, at=space, labels=NA,tck=-0.01, cex=.9)
text(x=space, y=rep(ypp,10), labels=top10all, adj=1, srt=90, xpd = NA)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="All 18S", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="F", cex=1) 
#####################################################################################

############################################################
# start column 2  metagenomes
############################################################
top10pat<-top10patMG
####    TARA
rr<-c(16) # column of response variable  !!!!
r<-1  ### number of levels
# tara ocean surface - comparable to mala
pdat<-dat_genome3 %>%   ## unique(dat11$Cruise.x)  names(dat_genome3) unique(pdat$filter_size) 
  filter(Cruise=="TARA") %>% 
  filter(Depth < 200) %>% # was Depth_region=="SRF" for 18s, hist(pdat$Depth)
  filter(filter_size_cat=="0.2-3" )  # .5-5 for 18s

j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Tara surface meta", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="G", cex=1)    
#mtext(text="b",side=1, line=-13.0,cex=cx, adj=.08, font=2)
############################################################
# no malaspina surface 2nd coloumn 2nd plot
frame()  
############################################################
# malaspina profile 2nd column  3rd plot
pdat<-dat_genome3 %>% 
  filter(Cruise=="Malaspina") %>% 
  filter(area=="prof") %>%    # unique(dat_genome3$area)
  filter(filter_size_cat=="0.2-3" ) 
 j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina profile meta", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="H", cex=1)  
############################################################
# no metagenomw RND 2nd coloumn 4th plot
frame()  
############################################################
###################################################################################
### metagenome malaspina deep  2nd column 5th plot
pdat<-dat_genome3 %>% 
  filter(Cruise=="Malaspina") %>% 
  filter(area=="deep") %>%    # unique(dat_genome3$area)
  filter(filter_size_cat=="0.2-3" ) 
j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
#  med<-tapply(resp,list(fac1),median,na.rm=T)
#  er<-tapply(resp,list(fac1),plotrix::std.error,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina deep meta", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="I", cex=1)  
############################################################
# ALL  metagenome 2nd col th plot  #
j<-rr
pdat<-dat_genome3  %>% 
  filter(filter_size_cat=="0.2-3" ) 

fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd =bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2)
axis(1, las=2, at=space, labels=NA,tck=-0.01, cex=.9)
text(x=space, y=rep(ypp,10), labels=top10all, adj=1, srt=90, xpd = NA)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="All metagenome", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="J", cex=1) 
#####################################################################################
#####################################################################################
#   start plot is line 287

#####################################################################################
###   mal ocean s- 3rd 1st three      unique(pdat$ocean)
pdat<-dat_genome3 %>% 
   filter(complete.cases(ocean))  %>% 
   mutate(ocean_basin=factor(ocean,levels=c("Atlantic",
                       "Pacific","Indian"))) %>% 
    mutate(ocean_basin = forcats::fct_recode(ocean_basin, "Indian Ocean"="Indian",
                               "Pacific Ocean"="Pacific","Atlantic Ocean"="Atlantic")) %>% 
  filter(filter_size_cat=="0.2-3" ) 

fac3<-factor(pdat$ocean_basin)  # ocean_basin or ocean2
ll<-levels(fac3)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-((pdat1[,j])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   # , log="y"
    if (i==r) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=F, tck=-0.01, cex=.9) #
      #text(x=space, y=rep(ypp,10), labels=top10pat, adj=1, srt=90, xpd = NA)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-ll
    text(tttlx,tttly, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("K","L","M")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
  }
}
#####################################################################################
###   lat zones,  oceans- 3rd column  last 3 plots
pdat<-dat_genome3 %>% # names(dat11)
  filter(complete.cases(Latitude_cat)) %>%  # levels(pdat$Latitude_cat)
  filter(filter_size_cat=="0.2-3" ) 

fac3<-factor(pdat$Latitude_cat)  # ocean_basin or ocean2
ll<-levels(fac3)
ll2<-levels(pdat$Latitude_cat)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-((pdat1[,j])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   # , log="y"
    if (i==r) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=F, tck=-0.01, cex=.9) #
      text(x=space, y=rep(ypp,10), labels=top10all, adj=1, srt=90, xpd = NA)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-ll2
    text(tttlx,tttly, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("N","O","P")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
  }
}


#####################################################################################
# at end add titlas
cqq<-.8
l11<-.2
mtext("Percent of reads per sample", 2,las=0,line=1.7, cex=cqq, at=.5 ,outer=TRUE)
mtext("Phylum", 1,las=0,line=6, cex=cqq, at=.5 ,outer=TRUE)

```




## boxplots genome
percents of phyla DNA from metagenome - fig 3 in manuscript
```{r barplots_gen}
lfil<-c(".5-5","5-20","20-200","180-2000")
n<-16
cc1<-rainbow(n)
cc1<-c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1','#35978f','#01665e','#003c30', "black")
space<-c(1:10) #
cc1<-cc1[1:length(space)]
yylim<-c(1,100)
tttlx<-11  ## location of plot label
tttly<-60
tttletx<-10    # location iof letter
tttlety<-90
ypp<-0.65
bxlwd<-1.5  # bos line width
mean_pch<-19 # pch of mean 
mean_cex<-1.3 #size o mean circle 
ylabat<-c(1,2,6,11,21,51,101)  # need to trick for y label - no 0s and added 1 below
ylab<-ylabat-1


############################################################
#### plot    names(dat1)
par(mfcol=c(4,3), mar=c(.2,1,1,2), oma=c(8,3,0,0), xpd = NA) 
########


############################################################
# start column 1  metagenomes
############################################################
top10pat<-top10patMG # top10patMG
####    TARA
rr<-c(16) # column of response variable  !!!!
r<-1  ### number of levels
# tara ocean surface - comparable to mala
pdat<-dat_genome3 %>%   ## unique(dat11$Cruise.x)  names(dat_genome3) unique(pdat$filter_size) 
  filter(Cruise=="TARA") %>% 
  filter(Depth < 200) %>% # was Depth_region=="SRF" for 18s, hist(pdat$Depth)
  filter(filter_size_cat=="0.2-3" )  # .5-5 for 18s

j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2, at=ylabat , labels= ylab)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Tara surface", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="A", cex=1)    
#mtext(text="b",side=1, line=-13.0,cex=cx, adj=.08, font=2)

############################################################
# malaspina profile 2nd column  3rd plot
pdat<-dat_genome3 %>% 
  filter(Cruise=="Malaspina") %>% 
  filter(area=="prof") %>%    # unique(dat_genome3$area)
  filter(filter_size_cat=="0.2-3" ) 
 j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2, at=ylabat , labels= ylab)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina profile", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="B", cex=1)  

############################################################
###################################################################################
### metagenome malaspina deep 
pdat<-dat_genome3 %>% 
  filter(Cruise=="Malaspina") %>% 
  filter(area=="deep") %>%    # unique(dat_genome3$area)
  filter(filter_size_cat=="0.2-3" ) 
j<-rr
fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2, at=ylabat , labels= ylab)
axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
#  med<-tapply(resp,list(fac1),median,na.rm=T)
#  er<-tapply(resp,list(fac1),plotrix::std.error,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="Malaspina deep", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="C", cex=1)  
############################################################
# ALL  metagenome th plot  #
j<-rr
pdat<-dat_genome3  %>% 
  filter(filter_size_cat=="0.2-3" ) 

fac1<-pdat$Phylum
resp<-((pdat[,j])*100)+1
b<-boxplot(resp~fac1, outline=F,
           col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd =bxlwd, log="y")   
box(lwd=1.2)
axis(2,las=2, at=ylabat , labels= ylab)
axis(1, las=2, at=space, labels=NA,tck=-0.01, cex=.9)
text(x=space, y=rep(ypp,10), labels=top10pat, adj=1, srt=90, xpd = NA)
#add means
q<-tapply(resp,list(fac1),mean,na.rm=T)
s=as.vector(q)
points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
text(tttlx,tttly, label="All metagenome", cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
text(tttletx,tttlety, label="D", cex=1) 
#####################################################################################
#####################################################################################
#   start plot is line 287

#####################################################################################
###   mal ocean s- 3rd 1st three      unique(pdat$ocean)
pdat<-dat_genome3 %>% 
   filter(complete.cases(ocean))  %>% 
   mutate(ocean_basin=factor(ocean,levels=c("Atlantic",
                       "Pacific","Indian"))) %>% 
    mutate(ocean_basin = forcats::fct_recode(ocean_basin, "Indian Ocean"="Indian",
                               "Pacific Ocean"="Pacific","Atlantic Ocean"="Atlantic")) %>% 
  filter(filter_size_cat=="0.2-3" ) 

fac3<-factor(pdat$ocean_basin)  # ocean_basin or ocean2
ll<-levels(fac3)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-((pdat1[,j])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   # , log="y"
    if (i==r) {
      box(lwd=1.2)
      axis(2,las=2, at=ylabat , labels= ylab)
      axis(1, las=2, at=space, labels=F, tck=-0.01, cex=.9) #
      text(x=space, y=rep(ypp,10), labels=top10pat, adj=1, srt=90, xpd = NA)
    }
    else {
      box()
      axis(2,las=2, at=ylabat , labels= ylab)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-ll
    text(tttlx,tttly, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("E","F","G")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
  }
}
########
frame()
#####################################################################################
###   lat zones,  oceans- 3rd column  last 3 plots
pdat<-dat_genome3 %>% # names(dat11)
  filter(complete.cases(Latitude_cat)) %>%  # levels(pdat$Latitude_cat)
  filter(filter_size_cat=="0.2-3" ) 

fac3<-factor(pdat$Latitude_cat)  # ocean_basin or ocean2
ll<-levels(fac3)
ll2<-levels(pdat$Latitude_cat)
r<-length(ll)
#  i<-1  ;j<-29
for (i in 1:r){  # filters
  for (j in rr){    #resp in rows
    pdat1<-pdat[fac3==ll[i],]  #  levels(dat1$Depth_cat)
    fac1<-pdat1$Phylum
    #fac2<-dat11$filter_crit_cat
    resp<-((pdat1[,j])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y")   # , log="y"
    if (i==r) {
      box(lwd=1.2)
      axis(2,las=2, at=ylabat , labels= ylab)
      axis(1, las=2, at=space, labels=F, tck=-0.01, cex=.9) #
      text(x=space, y=rep(ypp,10), labels=top10pat, adj=1, srt=90, xpd = NA)
    }
    else {
      box()
      axis(2,las=2, at=ylabat , labels= ylab)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)
    lln<-ll2
    text(tttlx,tttly, label=lln[i], cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("H","I","J")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
  }
}

#####################################################################################
# at end add titlas
cqq<-.8
l11<-.2
mtext("Percent of reads per sample", 2,las=0,line=1.7, cex=cqq, at=.5 ,outer=TRUE)
mtext("Phylum", 1,las=0,line=6, cex=cqq, at=.5 ,outer=TRUE)

```


## metagen depth plot
metagenome sequnces of phyla through depth
fig 2C in manucript
```{r depth_plots}
##################################
## get combined order of phym and reorder factors in dataframes
# levels(dat_genome3$Phylum)
# levels(dat11$Phylum)
top10pat_all<-c("Arthropoda","Cnidaria","Nematoda","Placozoa","Tardigrada","Chordata","Ctenophora","Annelida","Mollusca")
dat_genome_a<-dat_genome2 %>%   # unique(sum1$fac)  names(dat_genome_a) names(dat_genome2)
  filter(Phylum %in% top10pat_all) %>%   # limit to top ten phyla
  mutate(Phylum=factor(Phylum, levels=top10pat_all))  %>% #  top10pat_all or top10patMG- metagenome
  filter(Cruise=="Malaspina") %>% 
    filter(filter_size_cat=="0.2-3" ) %>% 
  filter(area=="prof") %>% # use both deep and profile!!!!!!!!!!!!
    # limit gene domain    !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(gene_domain=="ec4")  %>%   # unique(dat_genome2$gene_domain)
   # limit to target DMAP filter
  filter(pid=="PID_70") %>%  #  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(complete.cases(Depth_cat)) %>% 
  group_by(Depth_cat,Phylum) %>% 
  summarize(mean=mean(reads_percent,na.rm=T),median=median(reads_percent,na.rm=T)
            ,n(),se=plotrix::std.error(reads_percent))

########################################
### plot set up set up colors and factors
cc1<-c("black",'#543005','#8c510a','#bf812d','#dfc27d','#c7eae5','#80cdc1','#35978f','#01665e'
       ,'#003c30')
space<-c(1:9) # number of phylums !!
cc1<-cc1[1:length(space)]
fac3<-factor(dat_genome_a$Depth_cat)
ll<-levels(fac3)
#  get numbers for depth_cat
Dep_num <-c(5,110,500,1500,2500,4000)  # 
dep2<-data.frame(cbind(ll,Dep_num),stringsAsFactors = F)
names(dep2)[1]<-"Depth_cat"
dep2$Dep_num<-as.numeric(dep2$Dep_num)
# add color
depcol<-data.frame(cbind(top10pat_all,cc1),stringsAsFactors = F)
depcol$pch<-c(0:4,0:3)
names(depcol)[1]<-"Phylum"

yylim<-c(4057,0)
xxlim<-c(1,80)

##################################################################################################
###############################   start plotting    !!!!!!

##################################################################################################
###############################  plot
    par(mfcol=c(1,1),mar=c(0,.5,1,.5), oma=c(1,4,3,9)) # ,

##########################################################

##   begin set up for metagenome  plot
### at depth_num and color for plot
pdat<-dat_genome_a %>% 
    left_join(dep2) %>% 
    left_join(depcol)
pdat<-data.frame(pdat)
xxlim<-c(1,80)
rr<-3 # response in col  !!!!  reads_percent
r<-length(ll)
#  names(pdat)
pdat$xx<-pdat$Dep_num
pdat$xxx<-jitter(pdat$xx,2)
pdat$resp<-((pdat[,rr])*100)+1 
pdat$er<-((pdat$se)*100)+1 
pdat$cc<-scales::alpha(pdat$cc1,1)
pdat$cc2<-scales::alpha(pdat$cc1,0.4)
pdat$pch2<-pdat$pch
###  metagenome plot
    plot(pdat$xxx~pdat$resp, col=pdat$cc, axes=F,ylim=yylim,xlim=xxlim, pch=pdat$pch2,
         ylab="",xlab="",cex=1.5, lwd=2)   # log="x", 
      axis(2,las=2,tck=-0.05, labels = TRUE)
      axis(3, las=1,tck=-0.05, cex=.9)
# error bars
      error.bar(pdat$resp,pdat$xxx,pdat$er, col=pdat$cc2)
#  make lines to connect phylums   i=1
      for (i in 1:length(ll)){
        pdat2<-pdat[pdat$Phylum==top10pat[i],]
      lines(x=pdat2$resp,y=pdat2$xxx,col=pdat2$cc1,type = "l")
      }
# add letter
      # letter
      tttletx<--13
      tttlety<- -450
     text(tttletx,tttlety, label="C", cex=1) 
 ### add label and color for depth xone
     tcol<-c('#eff3ff','#bdd7e7','#6baed6','#3182bd','#08519c', "black")
     tlab<-c("Surface","Photic","Mesopelagic","Upper Bathypelagic", "Lower Bathypelagic", "Abyssopelagic")
     tgap<-c(-50,50,200,1000,2000,4000,4250)

     lef<-72 # 90 # lef,ri  norm, 75, 85 log  90,110
     ri<-83
     i<-1  ## updated
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-2
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-3
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-4
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",143), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), border=tcol[i], fill=tcol[i], col=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-5
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",230), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-6
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",21.5), cex=0.5, justify=c('l') , leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     #  y=top start, textlist-make number smaller and bottom will be shorter
     box(lwd=1.2)
    #text(tttlx,tttly, label=ll[i], cex=1)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))  
        
#######################################
## final additions
# legends
     legx<-85
      legend(x=legx, y= 0,legend=tlab, col=tcol, pch=15, bty='n',title="Depth zone",
            y.intersp = 1,xpd = NA)
      legend(x=legx, y= 1500,legend=depcol$Phylum, col=depcol$cc1, pch=depcol$pch, bty='n',
             title="Phylum", y.intersp = 1,xpd = NA)
    mtext("% of metazoan sequences",side=3,line=1.5,at= .5 , outer = TRUE)
    mtext("Depth (m)",side=2,line=3, outer = TRUE)

#####################################################################################
    #   save    7.36 and 6.65 pdf in genome plots
    
```

## relationships from benthos
check how metagenomes change wiht dist from depth
No clear relationship, not mentioned in manuscript
```{r depth_plots}
##################################
## get combined order of phym and reorder factors in dataframes
# levels(dat_genome3$Phylum)
# levels(dat11$Phylum)
top10pat_all<-c("Arthropoda","Cnidaria","Nematoda","Placozoa","Tardigrada","Chordata","Ctenophora","Annelida","Mollusca")
dat_genome_d<-dat_genome2 %>%   # unique(sum1$fac)  names(dat_genome_a) names(dat_genome2)
  filter(Phylum %in% top10patMG) %>%   # limit to top ten phyla
  mutate(Phylum=factor(Phylum, levels=top10patMG))  %>% 
    filter(filter_size_cat=="0.2-3" ) %>% 
    # limit gene domain    !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(gene_domain=="ec4")  %>%   # unique(dat_genome2$gene_domain)
   # limit to target DMAP filter
  filter(pid=="PID_70") %>%  #  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(complete.cases(Depth_cat)) %>% 
##  specific to this plot
  filter(Phylum=="Nematoda") %>%
  mutate(dist_bot=((global_depth)*-1) - Depth)


##################################################################################################
###############################   start plotting    !!!!!!

p<-ggplot(dat_genome_d, aes(x = dist_bot, y = reads_percent, size=Depth)) +
  geom_point() +
  geom_smooth(method='lm') +
  geom_smooth(method='loess') +
  # between facets
  facet_grid(~Cruise, space = "fixed", scales="free") +  # , scales = "fixed"

  labs(y = "Percent sequences", x = "Distance to benthos (m)")  +
  theme_bw()  + # classic -no grids or bw or minimal
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) + # 
  #theme(legend.position="none") +
  theme(panel.grid.major = element_blank())  # remove grids
  #theme(panel.spacing = unit(0.8, "lines"))  # increase margins


#  export
ggsave(filename="/Users/geraldn/Dropbox/Documents/KAUST/eDNA/DMAP/R/plots/comparison_2019/DMAP_dist_benthos_pf7_9plot_filt_0.2-3.pdf", plot = p)


    #   save    7.36 and 6.65 pdf in genome plots
    
```

## 2 depth plots
plot both 18s and metagenome DNA abudancne
now used in manuscript - added Nov 2020
```{r 2depth_plots}
##################################
## get combined order of phym and reorder factors in dataframes
# levels(dat_genome3$Phylum)
# levels(dat11$Phylum)
top10pat_all<-c("Arthropoda","Cnidaria","Nematoda","Placozoa","Tardigrada","Chordata","Ctenophora","Annelida","Mollusca")
dat_genome_a<-dat_genome2 %>%   # unique(sum1$fac)  names(dat_genome_a) names(dat_genome2)
  filter(Phylum %in% top10pat_all) %>%   # limit to top ten phyla
  mutate(Phylum=factor(Phylum, levels=top10pat_all))  %>% 
  filter(Cruise=="Malaspina") %>% 
    filter(filter_size_cat=="0.2-3" ) %>% 
  filter(area=="prof") %>% # use both deep and profile!!!!!!!!!!!!
    # limit gene domain    !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(gene_domain=="ec4")  %>%   # unique(dat_genome2$gene_domain)
   # limit to target DMAP filter
  filter(pid=="PID_70") %>%  #  !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(complete.cases(Depth_cat)) %>% 
  group_by(Depth_cat,Phylum) %>% 
  summarize(mean=mean(reads_percent,na.rm=T),median=median(reads_percent,na.rm=T)
            ,n(),se=plotrix::std.error(reads_percent))

##   18S  
dat11 = as.data.frame(dat11)

Amp_a <- dat11 %>% 
  filter(Sample.ID != "1001_DNA") %>% 
  filter(Phylum%in% top10pat_all) %>% 
  mutate(Phylum=factor(Phylum, levels=top10pat_all)) %>% 
  mutate(rich=as.numeric(rich))  %>% 
  filter(Cruise.x=="Malaspina") %>% 
  filter(DNAID=="DNA") %>% 
  filter(!(grepl("deep",Sample.ID) | grepl("M",Sample.ID))) %>% # filte !!!!!!! comparible to metagenome
  filter(complete.cases(Depth_cat)) %>% 
  group_by(Depth_cat,Phylum) %>% 
  summarize(mean=mean(new_prop_reads,na.rm=TRUE), median=median(new_prop_reads,na.rm=TRUE)
        ,n=n(), se=plotrix::std.error(new_prop_reads), rich_mean=mean(rich, na.rm=TRUE), rich_max=max(rich) , rich_se=plotrix::std.error(rich, na.rm=TRUE)  )
########################################    dat11$rich
### plot set up set up colors and factors
cc1<-c("black",'#543005','#8c510a','#bf812d','#dfc27d','#c7eae5','#80cdc1','#35978f','#01665e'
       ,'#003c30')
space<-c(1:9) # number of phylums !!
cc1<-cc1[1:length(space)]
fac3<-factor(Amp_a$Depth_cat)
ll<-levels(fac3)
#  get numbers for depth_cat
Dep_num <-c(5,110,500,1500,2500,4000)  # 
dep_labels<-c(0,1000,2000,3000,4000)
dep2<-data.frame(cbind(ll,Dep_num),stringsAsFactors = F)
names(dep2)[1]<-"Depth_cat"
dep2$Dep_num<-as.numeric(dep2$Dep_num)
# add color
depcol<-data.frame(cbind(top10pat_all,cc1),stringsAsFactors = F)
depcol$pch<-c(0:4,0:3)
names(depcol)[1]<-"Phylum"

yylim<-c(4070,0)

##################################################################################################
###############################   start plotting    !!!!!!

pdf(paste(dir, export_cor_plot,'depth_2_plots_18s_gen.pdf',sep = ""), width=11, height=9 )

##################################################################################################
###############################  plot
    par(mfcol=c(1,2),mar=c(0,.7,1,.5), oma=c(1,4,3,9), xpd=NA) # ,

##########################################################
##   begin set up for 18s plot
### at depth_num and color for plot
pdat<-Amp_a %>% 
    left_join(dep2) %>% 
    left_join(depcol)
pdat<-data.frame(pdat)

xxlim<-c(0,4.5)
rr<-7 # response in col  !!!!  rich   names(pdat)
r<-length(ll)
#  names(pdat)
pdat$xx<-pdat$Dep_num
pdat$xxx<-jitter(pdat$xx,2)
pdat$resp<-((pdat[,rr]))+0.1 
pdat$er<-((pdat$rich_se))+0.1
pdat$cc<-scales::alpha(pdat$cc1,1)
pdat$cc2<-scales::alpha(pdat$cc1,0.4)
pdat$pch2<-pdat$pch
### start    18S plot
    plot(pdat$xxx~pdat$resp, col=pdat$cc, axes=F, ylim=yylim, xlim=xxlim, pch=pdat$pch2,
         ylab="",xlab="",cex=2.5, lwd=3.8)   # log="x", 
      axis(2,las=2,tck=-0.025)
      axis(3, las=1,tck=-0.025, cex=.9)
# error bars
      error.bar(pdat$resp,pdat$xxx,pdat$er, col=pdat$cc2)
#  make lines to connect phylums   i=1
      for (i in 1:length(ll)){
        pdat2<-pdat[pdat$Phylum==top10pat_all[i],]
      lines(x=pdat2$resp,y=pdat2$xxx,col=pdat2$cc1,type = "l", lwd= 2.5)
      }

# letter
      tttletx<- -0.4
      tttlety<- -250
     text(tttletx,tttlety, label="A", cex=1) 
     #################################3
     ##    !!!!!!!   no longer used, only next one for second plot  !!!!###################3
 ### add label and color for depth xone
     tcol<-c('#eff3ff','#bdd7e7','#6baed6','#3182bd','#08519c', "black")
     tlab<-c("Surface","Photic","Mesopelagic","Upper Bathypelagic", "Lower Bathypelagic", "Abyssopelagic")
     tgap<-c(-50,50,200,1000,2000,4000,4250)

     lef<-75 # 90 # lef,ri  norm, 75, 85 log  90,110
     ri<-82.7
     i<-1  ## updated
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-2
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-3
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-4
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",143), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), border=tcol[i], fill=tcol[i], col=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-5
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",250), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-6
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",19), cex=0.5, justify=c('l') , leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     box(lwd=1.2)
    #text(tttlx,tttly, label=ll[i], cex=1)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
     
##########################################################
##   begin set up for metagenome  plot
### at depth_num and color for plot
pdat<-dat_genome_a %>% 
    left_join(dep2) %>% 
    left_join(depcol)
pdat<-data.frame(pdat)
xxlim<-c(1,80)
rr<-3 # response in col  !!!!  reads_percent
r<-length(ll)
#  names(pdat)
pdat$xx<-pdat$Dep_num
pdat$xxx<-jitter(pdat$xx,2)
pdat$resp<-((pdat[,rr])*100)+1 
pdat$er<-((pdat$se)*100)+1 
pdat$cc<-scales::alpha(pdat$cc1,1)
pdat$cc2<-scales::alpha(pdat$cc1,0.4)
pdat$pch2<-pdat$pch
###  metagenome plot
    plot(pdat$xxx~pdat$resp, col=pdat$cc, axes=F,ylim=yylim,xlim=xxlim, pch=pdat$pch2,
         ylab="",xlab="",cex=2.5, lwd=3.8)   # log="x", 
      axis(2,las=2,tck=-0.025, labels = c("","","","",""), at =dep_labels)  
      axis(3, las=1,tck=-0.025, cex=.9)
# error bars
      error.bar(pdat$resp,pdat$xxx,pdat$er, col=pdat$cc2)
#  make lines to connect phylums   i=1
      for (i in 1:length(ll)){
        pdat2<-pdat[pdat$Phylum==top10pat_all[i],]
      lines(x=pdat2$resp,y=pdat2$xxx,col=pdat2$cc1,type = "l", lwd=2.5)
      }
# add letter
     text(-5,tttlety, label="B", cex=1) 
 ### add label and color for depth xone
     tcol<-c('#eff3ff','#bdd7e7','#6baed6','#3182bd','#08519c', "black")
     tlab<-c("Surface","Photic","Mesopelagic","Upper Bathypelagic", "Lower Bathypelagic", "Abyssopelagic")
     tgap<-c(-50,0,200,1000,2000,4000,4250)

     lef<-75 # 90 # lef,ri  norm, 75, 85 log  90,110
     ri<-82.9
     i<-1  ## updated
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",110), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-2
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-3
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",113), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-4
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",143), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), border=tcol[i], fill=tcol[i], col=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-5
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",210), cex=0.5, justify=c('l'), leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     i<-6
     plotrix::textbox(x=c(lef,ri), y=tgap[c(i)], textlist=rep("L",20.5), cex=0.5, justify=c('l') , leading=0.5, box=T,
                      adj=c(0,0), col=tcol[i], border=tcol[i], fill=tcol[i], density=NULL,
                      angle=45, margin=0)
     box(lwd=1.2)
    #text(tttlx,tttly, label=ll[i], cex=1)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))  
        
#######################################
## final additions
# legends
     legx<-85
      legend(x=legx, y= 0,legend=tlab, col=tcol, pch=15, bty='n',title="Depth zone",
            y.intersp = 1,xpd = NA)
      legend(x=legx, y= 1000,legend=depcol$Phylum, col=depcol$cc1, pch=depcol$pch, bty='n',
             title="Phylum", y.intersp = 1,xpd = NA , pt.cex=1.5, pt.lwd=2.8)
    mtext("Richness (metabarcode)",side=3,line=1.5,at= .22 , outer = TRUE)
    mtext("Percent abundance (metagenome)",side=3,line=1.5,at= .77 , outer = TRUE)
    mtext("Depth (m)",side=2,line=3.0, outer = TRUE)

#####################################################################################
    dev.off()
    
    
    #   save     depth_18s_90_genec4_70_prof
    
```

# filter plots
plot 18s and genome by filter
changed 18s to richness
```{r filter_plot}


    cc1<-c('#543005','#8c510a','#bf812d','#dfc27d','#f6e8c3','#c7eae5','#80cdc1','#35978f','#01665e','#003c30', "black")
  cc1<-cc1[1:10]
    space<-c(1:10) #
    cc1<-cc1[1:length(space)]
    yylim<-c(1,100)
    tttlx<-11  ## location of plot label
    tttly<-35
    tttletx<-10.3    # location of letter
    tttlety<-90
    ypp<-0.65
    bxlwd<-1.5  # bos line width
    mean_pch<-19 # pch of mean 
    mean_cex<-1.3 #size o mean circle 
#######################################################################
#######################################################################
    pdf(paste(dir, export_cor_plot,'boxplot_filter_comp_rich_abund_phyl.pdf',sep = ""), width=11, height=8 )
    
    par(mfcol=c(3,3), mar=c(.2,2.5,1,2), oma=c(8,3,0,0), xpd = TRUE) 
    

##   tara 18s data, first column
pdat<-Amp_10 %>%   ## unique(dat11$Cruise.x)  names(dat1) unique(pdat$filter_size_cat) 
  filter(Cruise.x=="TARA") %>% 
  filter(!filter_size_cat=="5-20") %>%   # because half the samples of others
  mutate(filter_size_cat=factor(filter_size_cat, levels=c(".5-5","20-200","180-2000"))) %>% 
  group_by(Latitude,Longitude,Depth_region) %>% 
  mutate(filters_location=n()) %>% # 1872
  ungroup() %>% 
  filter(filters_location>20) # 1431 included to include places with 3 filter sizes
pdat<-data.frame(pdat)
# summary(pdat$filter_size_cat)    !!!!!!   53 locations   !!!!!!!!
    rr<-8 # column with resp variable  names(pdat)
    fac2<-pdat$filter_size_cat
    ll<-levels(fac2)
    lfil<-ll
    
## start plot loop
for (i in 1:length(ll)){  # filters
    pdat2<-pdat[pdat$filter_size_cat==ll[i],]
    fac1<-pdat2$Phylum
    resp<- (pdat2[,rr]) + 1  # add so axis can be log trans
  ## specify for different plots  
        if (i==1) {
      yylim2<-c(1,70)
      tttly2<-30
      tttlety2<-65
    }
    else {
     yylim2<-c(1,300)
      tttly2<-100
      tttlety2<-270
    }
    
  ## j plot  
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F, at=space ,border=cc1, ylim=yylim2, boxlwd = bxlwd, ann=FALSE, log="y")   # , log="y"  , 
    if (i==3) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=FALSE, tck=-0.01, cex=.9)
      ##   !! only when need x axis labels
      text(x=space, y=rep(ypp,10), labels=top10pat18s, adj=1, srt=60, xpd = NA)  # !!!!!
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)

    text(tttlx,tttly2, label=bquote( .(lfil[i])*mu*m), cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("A","B","C","D")
    text(tttletx,tttlety2, label=ttlet[i], cex=1)  
}
 
    
    #############3 
##   tara genome, second column
pdat<-dat_genome3 %>%   ## unique(dat11$Cruise.x)  names(dat_genome3) unique(pdat$filter_size_cat) 
  filter(Cruise=="TARA") %>% 
  mutate(filter_size_cat=factor(filter_size_cat)) %>% 
  filter(!filter_size_cat=="1-20") %>%   # because half the samples of others
  mutate(filter_size_cat=factor(filter_size_cat, levels=c("0.1-0.22","0.2-3"))) %>% 
  group_by(Latitude,Longitude,Depth) %>% 
  mutate(filters_location=n()) %>% # 
  ungroup() %>% 
  filter(filters_location==20) # 300     15 samples !!!!!!!!!
pdat<-data.frame(pdat)
# summary(pdat$filter_size_cat)
    rr<-16  # resp column
    fac2<-pdat$filter_size_cat
    ll<-levels(fac2)
    lfil<-ll # for text filter size
## start plot loop
for (i in 1:length(ll)){  # filters
    pdat2<-pdat[pdat$filter_size_cat==ll[i],]
    fac1<-pdat2$Phylum

    resp<-((pdat2[,rr])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y", ann=FALSE)   # , log="y"
    if (i==2) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=FALSE,tck=-0.01, cex=.9)
      text(x=space, y=rep(ypp,10), labels=top10patMG, adj=1, srt=60, xpd = NA)  # !!!!!
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)

    text(tttlx,tttly, label=bquote( .(lfil[i])*mu*m), cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("D","E")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
}
 
   
frame() 

 ##  Malaspina genome, second column
pdat<-dat_genome3 %>%   ## unique(dat11$Cruise.x)  names(dat_genome3) unique(pdat$filter_size_cat) 
  filter(Cruise=="Malaspina") %>% 
  mutate(filter_size_cat=factor(filter_size_cat)) %>% 
  mutate(filter_size_cat=factor(filter_size_cat, levels=c("0.2-3","1-20"))) %>% 
  group_by(Latitude,Longitude,Depth) %>% 
  mutate(filters_location=n()) %>% # 
  ungroup() %>% 
  filter(filters_location==20) # 740     !!!! 37 smaples
pdat<-data.frame(pdat)
# summary(pdat$filter_size_cat)
    rr<-16  # resp column
    fac2<-pdat$filter_size_cat
    ll<-levels(fac2)
    lfil<-ll # for text filter size
## start plot loop
for (i in 1:length(ll)){  # filters
    pdat2<-pdat[pdat$filter_size_cat==ll[i],]
    fac1<-pdat2$Phylum

    resp<-((pdat2[,rr])*100)+1
    b<-boxplot(resp~fac1, outline=F,
               col="white",axes=F,at=space,border=cc1,ylim=yylim,boxlwd = bxlwd, log="y", ann=FALSE)   # , log="y"
    if (i==2) {
      box(lwd=1.2)
      axis(2,las=2)
      axis(1, las=2, at=space, labels=FALSE, tck=-0.01, cex=.9)
      text(x=space, y=rep(ypp,10), labels=top10patMG, adj=1, srt=60, xpd = NA)
    }
    else {
      box()
      axis(2,las=2)
      axis(1, las=1, padj=-1, at=space, labels=F,tck=-0.01, cex=.9)
    }
    #add means
    q<-tapply(resp,list(fac1),mean,na.rm=T)
    s=as.vector(q)
    points(space,s,pch=mean_pch,col=cc1,cex=mean_cex)

    text(tttlx,tttly, label=bquote( .(lfil[i])*mu*m), cex=1, pos=2)    #expression(mu * "ml")   bquote("Significance level is" ~ alpha == .(obj$alpha))
    ttlet<-c("F","G")
    text(tttletx,tttlety, label=ttlet[i], cex=1)  
}
    
    
    
      
    cqq<-.8
    l11<-.2
    mtext("Richness (metabarcode)", 2,las=0,line=0.4, cex=cqq, at=.5 ,outer=TRUE)
    mtext("Percent abundance (metagenome)", 2,las=0,line=-26.5, cex=cqq, at=.65 ,outer=TRUE)
    mtext("Percent abundance (metagenome)", 2,las=0,line=-53.5, cex=cqq, at=.65 ,outer=TRUE)
    mtext("Phylum", 1,las=0,line=4, cex=cqq, at=.5 ,outer=TRUE) 
    
    #  !!!!!!!!!!!!!
dev.off()


```


